Tropical Mathematics and the Lambda-calculus I

### 2023-11-30 CSL24/section4.tex
commit

As seen in Section \ref{section2}, tropical polynomials are piecewise-linear functions, hence concave and Lipschitz-continuous. Moreover, tps in finitely many variables are locally equivalent to tropical polynomials (except at some singular points), and are thus also concave and locally Lipschitz-continuous.
In this section we show that much of these properties extend also to tps with infinitely many variables, as those arising from the tropical relational model, %establish some topological and metric properties of tps, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
and by a tropical polynomial or power series we mean one with possibly infintely many variables.
The literature on tropical power series is often recent (e.g.~\cite{Porzio2021}), and several results we prove in this section are, to our knowledge, new.
%
%Remark that the map $(\_)^!$ is a surjective homomorphism (w.r.t.\ coKleisli composition and identity and function composition and identiy) on its image. {\color{red}In Ehrhard-Danos fanno vedere (Lemma 14) che \`e anche iniettivo. Non riesco a capire se anche a noi lo \`e, ma non vedo perch\'e non dovrebbe.}
%Remark that, since $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}\in\HOM{\LREL_!}{\mathlarger{\&}_{i=1}^n \model{A_i}}{\model B}\simeq\Lawv^{\prod_{i=1}^n ! \model {A_i}\times \model B}$, the tps $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}^!:\Lawv^{\mathlarger{+}_{i=1}^n \model{A_i}}\to\Lawv^{\model B} \simeq \prod_{i=1}^n\Lawv^{\model{A_i}}\to\Lawv^{\model B}$ has shape $t^!(x^1,\dots,x^n)_b=\inf_{\mu_i\in ! \model{A_i}}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Same for $\STDLC$.
%
%\begin{remark}
%The interpretations of $\BSTLC$-terms are matrices $t\in\HOM{\LREL}{\bigotimes_{i=1}^n !_{n_i}X_i}{Y}=\Lawv^{\prod_{i=1}^n !_{n_i}X_i\times Y}$.
%In such situation, we define $t^!:\Lawv^{\mathlarger{+}_{i=1}^n X_{i}}\to\Lawv^Y \simeq \prod_{i=1}^n\Lawv^{X_{i}}\to\Lawv^Y$ as $t^!(x^1,\dots,x^n)_b:=\inf_{\mu_i\in !_{n_i} X_i}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Clearly $t^!={\widetilde t}^!$, where $\widetilde t\in\Lawv^{\prod_{i=1}^n ! {X_i}\times Y}$ is the matrix $\widetilde t_{(\mu_1,\dots,\mu_n),b}:=t_{(\mu_1,\dots,\mu_n),b}$ if $\mu_i\in !_{n_i} X_i$ for all $i$, and $:=\infty$ otherwise, which has the same support as $t$.
%So, such $t^!$'s are special cases of tps and thus we can \emph{Wlog} only consider tps's.
%\end{remark}

@@

Looking at Fig~\ref{fig:plot1}, we see that $\varphi$, just like the polynomials $\varphi_{n}$, is non-decreasing and concave.
This is indeed always the case:

@@

\begin{proposition}\label{prop:troplinear}
\begin{enumerate}
\item If a tps $f: \overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$ arises from a matrix $\matr f:X\times Y\to \overline{\BB R}_{\geq 0}$ (i.e.~it is tropical linear), then $f$ is non-expansive (i.e.\ $1$-Lipschitz).  
\item If  $f: \overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$ arises from a matrix $\matr f: \ !_{n}X\times Y\to \overline{\BB R}_{\geq 0}$, then $f$ is $n$-Lipschitz-continuous.
\end{enumerate}
\end{proposition}
\begin{proof}[Proof sketch]
1). Using the fact that $f( x)_{b}= \inf_{a\in X}\{\matr f_{a,b}+ x_{a}\}$,
the problem reduces to: $|(\matr f_{a,b}- x_{a})- (\matr f_{a,b}- y_{a})| = |x_{a}- y_{a}|\leq \norm{x- y}_{\infty}$. %(see the Appendix).

@@

For any tropical polynomial $\varphi:\overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$, the associated matrix has shape $!_{\mathrm{deg}(\varphi)}(X)\times Y\to \overline{\BB R}_{\geq 0}$ (as a monomial $\mu_ix+c_{i}$ yields a matrix entry on $!_{\#\mu_i}X\times Y$). Hence, using Proposition \ref{prop:troplinear} 2., we have:
\begin{corollary}\label{prop:polylip}
Any tropical polynomial $\varphi:\overline{\BB R}_{\geq 0}^{X}\to\overline{\BB R}_{\geq 0}$ is $\mathrm{deg}(\varphi)$-Lipschitz continuous.
\end{corollary}

@@

\begin{theorem}\label{thmTLSlocLip}
%Let $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $d$ any metric.
%If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
%Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
%Therefore, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, %and all tps are non-negative, 
%we immediately have that 
All tps $f:\overline{\BB R}_{\geq 0}^X\to\overline{\BB R}_{\geq 0}$ are locally Lipschitz on $(0,\infty)^X$.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} f$.
\end{theorem}
%\begin{proof}
%By observing that $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, we apply the result that for all $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$ concave and locally bounded, with $V$ open and convex and $d$ any metric, $f$ is locally Lipschitz, with the stated Lipschitz constant on $\overline{B_{\delta}(x)}$. %
% Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
%\end{proof}
This crucial ingredient of the proof of this result is the following \autoref{th:locLip}, from which \autoref{thmTLSlocLip} immediately follows, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$ and all tps are non-negative.

\begin{wrapfigure}{r}{0.5\textwidth}%\begin{figure}
\begin{tikzpicture}[thick,scale=0.9]
    \coordinate (u) at (-2,0);
    \coordinate (v) at (2,0);
    \coordinate (y) at (-1.2,0);
    \coordinate (z) at (0,0);

@@

    \fill   (z) circle[radius=0pt]  node [xshift=-1cm, yshift=2.2cm] {$\overline{B_{2\delta}(z)}$};

@@

    \draw   (x) circle[radius=1cm] node [above] {$x$};
    \fill   (x) circle[radius=2pt]  node [xshift=10mm, yshift=10mm] {$\overline{B_{\delta}(x)}$};

    \draw   (x) circle[radius=3cm] node {};
    \fill   (x) circle[radius=0pt]  node [xshift=-1.7cm, yshift=3cm] {$\overline{B_{3\delta}(x)}$};
\end{tikzpicture}
\caption{Drawing of the proof of \autoref{th:locLip}.}
\label{fig:proof_loc_lip}
\end{wrapfigure} %%

\begin{theorem}\label{th:locLip}
Let $f:V\subseteq (\BB R^X,\norm\cdot) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $\norm\cdot$ any norm.
If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
\end{theorem}
\begin{proof}
 Call $\overline{B_{\delta}(x)}:=B_1$, $\overline{B_{3\delta}(x)}:=B_3$.
 It suffices to show that for all $x\in V$, there is $\delta>0$ s.t.\ $B_3\subseteq \mathrm{interior}(V)$, $K:=\max_{B_3}  \absv f$ exists and $f$ is $\frac{K}{\delta}$-Lipschitz on $B_1$.
A $\delta$ satisfying the first two conditions exists since $V$ is open and bcause $f$ is locally bounded and $B_3$ is compact.
We show that the third condition holds for all such $\delta$.
 For that, fix $y,z\in B_1$ and call $r:=\frac{d(y,z)}{2\delta}\in[0,1]$.
 We want to show that $\absv{f(y)-f(z)}\leq \frac{K}{\delta}d(y,z)=2Kr$.
 Wlog $y\neq z$, otherwise there is nothing to prove.

>>>>

As seen in Section \ref{section2}, tropical polynomials are piecewise-linear functions, hence concave and Lipschitz-continuous. Moreover, tps in finitely many variables are locally equivalent to tropical polynomials (except at some singular points), and are thus also concave and locally Lipschitz-continuous.
In this section we show that much of these properties extend also to tps with infinitely many variables, as those arising from the tropical relational model. %establish some topological and metric properties of tps, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
%and by a tropical polynomial or power series we mean one with possibly infintely many variables.
The literature on tropical power series is often recent (e.g.~\cite{Porzio2021}), and several results we prove in this section are, to our knowledge, new.
%
%Remark that the map $(\_)^!$ is a surjective homomorphism (w.r.t.\ coKleisli composition and identity and function composition and identiy) on its image. {\color{red}In Ehrhard-Danos fanno vedere (Lemma 14) che \`e anche iniettivo. Non riesco a capire se anche a noi lo \`e, ma non vedo perch\'e non dovrebbe.}
%Remark that, since $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}\in\HOM{\LREL_!}{\mathlarger{\&}_{i=1}^n \model{A_i}}{\model B}\simeq\Lawv^{\prod_{i=1}^n ! \model {A_i}\times \model B}$, the tps $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}^!:\Lawv^{\mathlarger{+}_{i=1}^n \model{A_i}}\to\Lawv^{\model B} \simeq \prod_{i=1}^n\Lawv^{\model{A_i}}\to\Lawv^{\model B}$ has shape $t^!(x^1,\dots,x^n)_b=\inf_{\mu_i\in ! \model{A_i}}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Same for $\STDLC$.
%
%\begin{remark}
%The interpretations of $\BSTLC$-terms are matrices $t\in\HOM{\LREL}{\bigotimes_{i=1}^n !_{n_i}X_i}{Y}=\Lawv^{\prod_{i=1}^n !_{n_i}X_i\times Y}$.
%In such situation, we define $t^!:\Lawv^{\mathlarger{+}_{i=1}^n X_{i}}\to\Lawv^Y \simeq \prod_{i=1}^n\Lawv^{X_{i}}\to\Lawv^Y$ as $t^!(x^1,\dots,x^n)_b:=\inf_{\mu_i\in !_{n_i} X_i}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Clearly $t^!={\widetilde t}^!$, where $\widetilde t\in\Lawv^{\prod_{i=1}^n ! {X_i}\times Y}$ is the matrix $\widetilde t_{(\mu_1,\dots,\mu_n),b}:=t_{(\mu_1,\dots,\mu_n),b}$ if $\mu_i\in !_{n_i} X_i$ for all $i$, and $:=\infty$ otherwise, which has the same support as $t$.
%So, such $t^!$'s are special cases of tps and thus we can \emph{Wlog} only consider tps's.
%\end{remark}

@@

Looking at Fig.~\ref{fig:plot1}, we see that $\varphi$, just like the polynomials $\varphi_{n}$, is non-decreasing and concave.
This is indeed always the case:

@@

\begin{proposition}\label{prop:troplinear}
\begin{enumerate}
\item If a tps $f: \overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$ arises from a matrix $\matr f:X\times Y\to \overline{\BB R}_{\geq 0}$ (i.e.~it is linear), then $f$ is non-expansive (i.e.\ $1$-Lipschitz).  
\item If  $f: \overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$ arises from a matrix $\matr f: \ !_{n}X\times Y\to \overline{\BB R}_{\geq 0}$, then $f$ is $n$-Lipschitz-continuous.
\end{enumerate}
\end{proposition}
\begin{proof}[Proof sketch]
1). Using the fact that $f( x)_{b}= \inf_{a\in X}\{\matr f_{a,b}+ x_{a}\}$,
the problem reduces to: $|(\matr f_{a,b}- x_{a})- (\matr f_{a,b}- y_{a})| = |x_{a}- y_{a}|\leq \norm{x- y}_{\infty}$. %(see the Appendix).

@@

When a tps is expressed as an $\inf$ of finitely many monomials we simply call it a tropical polynomial. 
For any such tropical polynomial $\varphi:\overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$, the associated matrix has shape $!_{\mathrm{deg}(\varphi)}(X)\times Y\to \overline{\BB R}_{\geq 0}$ (as a monomial $\mu_ix+c_{i}$ yields a matrix entry on $!_{\#\mu_i}X\times Y$). Hence, using Proposition \ref{prop:troplinear} 2., we have:
\begin{corollary}\label{prop:polylip}
Any tropical polynomial $\varphi:\overline{\BB R}_{\geq 0}^{X}\to\overline{\BB R}_{\geq 0}$ is $\mathrm{deg}(\varphi)$-Lipschitz continuous.
\end{corollary}

@@

\begin{theorem}\label{thmTLSlocLip}
%Let $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $d$ any metric.
%If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
%Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
%Therefore, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, %and all tps are non-negative, 
%we immediately have that 
All tps $f:\overline{\BB R}_{\geq 0}^X\to\overline{\BB R}_{\geq 0}$ are locally Lipschitz on $(0,\infty)^X$.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} f$.
\end{theorem}
%\begin{proof}
%By observing that $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, we apply the result that for all $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$ concave and locally bounded, with $V$ open and convex and $d$ any metric, $f$ is locally Lipschitz, with the stated Lipschitz constant on $\overline{B_{\delta}(x)}$. %
% Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
%\end{proof}
The fundamental ingredient of the proof of this result is the following \autoref{th:locLip}, from which \autoref{thmTLSlocLip} immediately follows, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$ and all tps are non-negative.

 %%

% \begin{wrapfigure}{r}{0.5\textwidth}%\begin{figure}
\begin{figure}
\begin{center}
\begin{tikzpicture}[thick,scale=0.7]
    \coordinate (u) at (-2,0);
    \coordinate (v) at (2,0);
    \coordinate (y) at (-1.2,0);
    \coordinate (z) at (0,0);

@@

    \fill   (z) circle[radius=0pt]  node [xshift=-1cm, yshift=1.6cm] {$\overline{B_{2\delta}(z)}$};

@@

    \draw   (x) circle[radius=1cm] node [above] {$x$};
    \fill   (x) circle[radius=2pt]  node [xshift=10mm, yshift=7mm] {$\overline{B_{\delta}(x)}$};

    \draw   (x) circle[radius=3cm] node {};
    \fill   (x) circle[radius=0pt]  node [xshift=-1.7cm, yshift=2cm] {$\overline{B_{3\delta}(x)}$};
\end{tikzpicture}
\end{center}
\caption{Drawing of the proof of \autoref{th:locLip}.}
\label{fig:proof_loc_lip}
\end{figure}
%\end{wrapfigure}

\begin{theorem}\label{th:locLip}
Let $f:V\subseteq (\BB R^X,\norm\cdot) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $\norm\cdot$ any norm.
If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
\end{theorem}
\begin{proof}
 Call $\overline{B_{\delta}(x)}:=B_1$, $\overline{B_{3\delta}(x)}:=B_3$.
 It suffices to show that for all $x\in V$, there is $\delta>0$ s.t.\ $B_3\subseteq \mathrm{interior}(V)$, $K:=\max_{B_3}  \absv f$ exists and $f$ is $\frac{K}{\delta}$-Lipschitz on $B_1$.
A $\delta$ satisfying the first two conditions exists since $V$ is open and because $f$ is locally bounded and $B_3$ is compact.
We show that the third condition holds for all such $\delta$.
 For that, fix $y,z\in B_1$ and call $r:=\frac{d(y,z)}{2\delta}\in[0,1]$.
 We want to show that $\absv{f(y)-f(z)}\leq \frac{K}{\delta}d(y,z)=2Kr$.
 Wlog $y\neq z$, otherwise there is nothing to prove.

### 2023-11-30 CSL24/section5bis.tex
commit

In many situations (e.g.~when dealing with computationally difficult problems) one does not look for algorithms to compute a function \emph{exactly}, but rather to approximate it (in an efficient way) within some error bound. In other common situations (e.g.~in differential privacy \cite{Alvim2011, Reed2010}) one needs to verify that an algorithm is not \emph{too sensitive} to errors, that is, that a small error in the input will produce a comparably small error in the output. 
In all these cases, it is common to consider forms of denotational semantics in which types are endowed with a \emph{behavioral metric}, that is, a metric on programs which accounts for differences in behavior. 
A fundamental insight coming from this line of work is that, if one can somehow   \emph{bound} the number of times that a program may duplicate its input, the resulting program will be \emph{Lipschitz-continuous}:  
if $M$ may duplicate at most $L$ times, then an error $\epsilon$ between two inputs will result in an error less or equal to $L\cdot \epsilon$ in the corresponding outputs \cite{Reed2010, Gaboardi2017}.
For instance, the higher-order program $M=\lambda f.\lambda x.f(f(x))$, which duplicates the functional input $f$, yields a $2$-Lipschitz map between the metric space $\BB R\multimap \BB R$ of non-expansive real functions and itself: if $f,g$ are two non-expansive maps differing by at most $\epsilon$ (i.e.~for which $|f(x)-g(x)|\leq \epsilon$ holds for all $x\in \BB R$), then the application of $M$ to $f$ and $g$ will produce two maps differing by at most $2\epsilon$.

@@

Yet, what about the good old, ``unbounded'', simply typed $\lambda$-calculus? Actually, by using unbounded duplications, one might lose the Lipschitz property. For instance, while the functions $M_{k}=\lambda x. k\cdot x: \BB R\to \BB R$ are all Lipschitz-continuous, with Lipschitz constant $k$, the function $M=\lambda x.x^{2}$ obtained by ``duplicating'' $x$ is not Lipschitz anymore: $M$ is, so to say, \emph{too} sensitive to errors. 
More abstractly, it is well-known that the category $\Met$ is \emph{not} cartesian closed, so it is not a model of $\STLC$ (yet, several cartesian closed \emph{sub-}categories of $\Met$ do exist, see e.g.~\cite{Clementino2006, PistoneFSCD2022}).
Still, one might observe that the program $M$ above is actually Lipschitz-continuous, if not globally, at least \emph{locally} (i.e.~over any compact set). Indeed, some cartesian closed categories of locally Lipschitz maps have been produced in the literature \cite{Ehrhard2011, PistoneLICS}, and a new example will be exhibited in this paper.

>>>>

In many situations (e.g.~when dealing with computationally difficult problems) one does not look for algorithms to compute a function \emph{exactly}, but rather to approximate it (in an efficient way) within some error bound. In other common situations (e.g.~in differential privacy \cite{Alvim2011, Reed2010}) one needs to verify that an algorithm is not \emph{too sensitive} to errors, that is, that a small error in the input will produce a comparably small error in the output. 
In all these cases, it is common to consider forms of denotational semantics in which types are endowed with a \emph{behavioral metric}, that is, a metric on programs which accounts for differences in behavior. 
A fundamental insight coming from this line of work is that, if one can somehow   \emph{bound} the number of times that a program may duplicate its input, the resulting program will be \emph{Lipschitz-continuous}:  
if $M$ may duplicate at most $L$ times, then an error $\epsilon$ between two inputs will result in an error less or equal to $L\cdot \epsilon$ in the corresponding outputs \cite{Reed2010, Gaboardi2017} (yet, this property may fail in a concurrent setting, see e.g.~\cite{Gebler2018}).
For instance, the higher-order program $M=\lambda f.\lambda x.f(f(x))$, which duplicates the functional input $f$, yields a $2$-Lipschitz map between the metric space $\BB R\multimap \BB R$ of non-expansive real functions and itself: if $f,g$ are two non-expansive maps differing by at most $\epsilon$ (i.e.~for which $|f(x)-g(x)|\leq \epsilon$ holds for all $x\in \BB R$), then the application of $M$ to $f$ and $g$ will produce two maps differing by at most $2\epsilon$.

@@

Yet, what about the good old, ``unbounded'', simply typed $\lambda$-calculus? Actually, by using unbounded duplications, one might lose the Lipschitz property. For instance, while the functions $M_{k}=\lambda x. k\cdot x: \BB R\to \BB R$ are all Lipschitz-continuous, with Lipschitz constant $k$, the function $M=\lambda x.x^{2}$ obtained by ``duplicating'' $x$ is not Lipschitz anymore: $M$ is, so to say, \emph{too} sensitive to errors. 
More abstractly, it is well-known that the category $\Met$ of metric spaces and non-expansive maps, is \emph{not} cartesian closed, so it is not a model of $\STLC$ (yet, several cartesian closed \emph{sub-}categories of $\Met$ do exist, see e.g.~\cite{Clementino2006, PistoneFSCD2022}).
Still, one might observe that the program $M$ above is actually Lipschitz-continuous, if not globally, at least \emph{locally} (i.e.~over any compact set). Indeed, some cartesian closed categories of locally Lipschitz maps have been produced in the literature \cite{Ehrhard2011, PistoneLICS}, and a new example will be exhibited in this paper.

### 2023-11-30 CSL24/sectionTaylip.tex
commit

\begin{corollary}\label{cor:taylor}
Suppose $x: A \vdash_{\STLC}M:B$ and $\vdash_{\STLC} N:A$. 
Then for all $t\in \Te{M}$ such that $\model{t}^{!}(\model{N})\neq \infty$, and $\delta>0$, the tps $\model{ x:A\vdash_{\STLC}M:B}^{!}$ is $\frac{\model{t}^{!}(\model{N}+2\delta)}{\delta}$-Lipschitz over the open ball
 $B_{\delta}(\model{N})$.
\end{corollary}
\begin{proof}
Thm.~\ref{thmTLSlocLip} yields the estimate $\max_{\overline{B_{3\delta}(\model{N})}}\model{M}^{!}$. As from Thm.~\ref{thm:taylor} 4.~it follows that $\model{t}^{!}\geq \model{M}^{!}$, we deduce that $K=\max_{\overline{B_{3\delta}(\model{N})}}\model{t}^{!}\geq \max_{\overline{B_{3\delta}(\model{N})}}\model{M}^{!}$ is also a local Lipschitz constant for $\model{M}^{!}$. Moreover, since $\model{t}^{!}$ is concave and non-decreasing, the $\max$ of $\model{t}^{!}$ is attained at the maximum point of $\overline{B_{3\delta}(\model{N})}$, that is, 
$K= \model{t}^{!}(x+3\delta)$. Finally, from $\model{t}^{!}(\model{N})<\infty$ and the continuity of $\model{t}^{!}$ we deduce $K<\infty$.
\end{proof}
%\begin{proof}[Proof sketch.]
%We apply the Lipschitz estimate from Thm.~\ref{thmTLSlocLip} for $\model{M}^{!}$,  and use the fact that, from Thm.~\ref{thm:taylor} 4.~it follows that $\model{t}^{!}\geq \model{M}^{!}$.
%To conclude, since $\model{t}^{!}$ is concave and non-decreasing, $\max_{\overline{B_{3\delta}(x)}}\model{t}^{!}= \model{t}^{!}(x+3\delta)$, which yields the stated Lipschitz estimate. 
%\end{proof}
%{\color{red}ADD PROOF SKETCH}

>>>>

\begin{corollary}\label{cor:taylor}
Suppose $x: A \vdash_{\STLC}M:B$ and $\vdash_{\STLC} N:A$. 
Then for all $t\in \Te{M}$ such that $\model{t}^{!}(\model{N})\neq \infty$, and $\delta>0$, the tps $\model{ x:A\vdash_{\STLC}M:B}^{!}$ is $\frac{\model{t}^{!}(\model{N}+3\delta)}{\delta}$-Lipschitz over the open ball
 $B_{\delta}(\model{N})$.
\end{corollary}
\begin{proof}
Thm.~\ref{thmTLSlocLip} yields the estimate $\max_{\overline{B_{3\delta}(\model{N})}}\model{M}^{!}$. As from Thm.~\ref{thm:taylor} 4.~it follows that $\model{t}^{!}\geq \model{M}^{!}$, we deduce that $K=\max_{\overline{B_{3\delta}(\model{N})}}\model{t}^{!}\geq \max_{\overline{B_{3\delta}(\model{N})}}\model{M}^{!}$ is also a local Lipschitz constant for $\model{M}^{!}$. Moreover, since $\model{t}^{!}$ is concave and non-decreasing, the $\max$ of $\model{t}^{!}$ is attained at the maximum point of $\overline{B_{3\delta}(\model{N})}$, that is, 
$K= \model{t}^{!}(x+3\delta)$. Finally, from $\model{t}^{!}(\model{N})<\infty$ and the continuity of $\model{t}^{!}$ we deduce $K<\infty$.
\end{proof}
%\begin{proof}[Proof sketch.]
%We apply the Lipschitz estimate from Thm.~\ref{thmTLSlocLip} for $\model{M}^{!}$,  and use the fact that, from Thm.~\ref{thm:taylor} 4.~it follows that $\model{t}^{!}\geq \model{M}^{!}$.
%To conclude, since $\model{t}^{!}$ is concave and non-decreasing, $\max_{\overline{B_{3\delta}(x)}}\model{t}^{!}= \model{t}^{!}(x+3\delta)$, which yields the stated Lipschitz estimate. 
%\end{proof}
%{\color{red}ADD PROOF SKETCH}
